---
title: "Epidemiological Report - COVID-19"
author: "Hugo Soubrier"
date: "``r Sys.Date()``"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
---

```{r set up, include = F}

knitr::opts_chunk$set(echo = FALSE,       
                      error = TRUE,
                      message = FALSE, 
                      warning = F)

```

```{r library loading}
#spatial analysis 
library(shiny)
library(sf)
library(ggsn)
library(raster)
library(leaflet)

#epidemiological analysis 
library(shinyWidgets)
library(sitrep)
library(skimr)
library(epikit)
library(incidence)
library(linelist)
library(magrittr)
library(utils)
library(lubridate)
library(aweek)
library(naniar)
library(ggplot2)
library(devtools)
library(plotly)
library(shinydashboard)
library(shinyjs)
library(janitor)
library(lubridate)
library(sunburstR)
library(knitr)
library(kableExtra)
library(rio)
library(here)
library(tidyselect)
library(tidyverse)
library(highcharter)
```
## Introduction 

```{r data loading & cleaning,}
#read data
data <- import(here("data", "covid_example_data.xlsx"))

# set data classes -----------------------------------

#create vector for the factors variables
fct_vars <- data %>% select(contains("sym"), #select all symptoms var
                            -sym_startdt_FALSE, #exclude the symptoms date var 
                            - sym_resolveddt_FALSE, #exclude the symptoms resolved date var 
                            case_gender, 
                            case_race, 
                            case_eth, 
                            hospitalized, 
                            died, 
                            died_covid, 
                            confirmed_case, 
                            covid_dx) %>% names()

#create vector for the date variables
date_vars <- data %>% select(case_dob_FALSE, contains("dt")) %>% names()

#create vector for numeric variables
numeric_vars <- data %>% select(case_zip, case_age, latitude_JITT, longitude_JITT) %>% names()

#create vector for character variables
character_vars <- data %>% select(PID) %>% names()


#apply data class to each vectors -
clean_data <- data %>% mutate(across(fct_vars, as_factor),
                across(date_vars, as_date), 
                across(numeric_vars, as.numeric), 
                across(character_vars, as.character))

# Tidy data -------------------
#create vector of var that need cleaning 
needtidy_vars <- clean_data %>% select(where(is.factor)) %>% colnames()

# use linelist::clean_data on it 
clean_data[needtidy_vars] <- clean_data %>% select(all_of(needtidy_vars)) %>% clean_data()

#Age 
#make NA all obs that have an age < 0
clean_data <- clean_data %>% mutate(case_age = ifelse(case_age<0, NA, case_age))

```

```{r spatial data loading, include = F}
#read spatial data and make the ZipCode variable the same type as the case_zip in the Epi data for joining - numeric
vector <- read_sf("data/vectors/FultonCountyZipCodes.shp") %>% 
  mutate(ZipCode = as.numeric(ZipCode))

#set the CRS to 4326 - WGS 1984
vector <- st_transform(vector, 4326)

#extract population data to perform aggregation - remove geometry for now
pop_data <- vector %>% 
  dplyr::select(ZipCode, Population) %>% 
  st_drop_geometry()

```

```{r data recoding & classifying, include = F}
# create age_cat var 
clean_data <- clean_data %>%  mutate(age_cat = case_when( case_age < 10 ~ "< 10", 
                                                between(case_age, 10, 20) ~ "10 - 20", 
                                                between(case_age, 20, 30) ~ "20 - 30", 
                                                between(case_age, 30, 40) ~ "30 - 40", 
                                                between(case_age, 40, 50) ~ "40 - 50", 
                                                between(case_age, 50, 60) ~ "50 - 60", 
                                                between(case_age, 60, 70) ~ "60 - 70", 
                                                case_age > 70 ~ "70 +"), 
                                     age_cat = if_else(is.na(age_cat), "unknown", age_cat), 
                                     age_cat = fct_relevel(age_cat, c("< 10", "10 - 20", "20 - 30", "30 - 40", "40 - 50", "50 - 60", "60 - 70", "70 +"))) %>% 
  
  relocate(age_cat, .after = case_age)

#recode all variables - if NA then set as unknown  

clean_data <- clean_data %>% mutate(across(fct_vars, as.character),
                      across(fct_vars, ~if_else(is.na(.x), "unknown", .x)),
                      across(fct_vars, as_factor))

#recode specific variables

clean_data <- clean_data %>% 
  mutate( hospitalized = recode(hospitalized, 
                                "yes" = "hospitalised",
                                "no" = "not hospitalised"))

clean_data <- clean_data %>% 
  mutate(died_covid = recode( died_covid, 
                              "under_review" = "under review"))


```

```{r incidence calculation, include = F}
#calculate incidence and get ISO-weeks for the outbreak for later uses
inci <- incidence(clean_data$pos_sampledt_FALSE, interval = "monday week")

```

## **Overview** 
Between the  `r min(inci$dates)` and the `r max(inci$dates)` (a period `r inci$timespan ` days) a total of `r inci$n` cases have been identified, of which `r fmt_count(clean_data, covid_dx == "confirmed")` are **confirmed** cases. 

- In Fulton County, and over this period, the **attack rate** is of `r fmt_ci_df(attack_rate(inci$n, population = sum(pop_data$Population), multiplier = 10000), percent = F)` per 10 000.

- Amongst all cases, at least `r fmt_count(clean_data, hospitalized == "hospitalised")` have been **hospitalised**, but `r fmt_count(clean_data, hospitalized == "unknown")` still have an **unknown** status 

- Using only cases with known outcome, the **Case Fatality Ratio** is of `r fmt_ci_df(case_fatality_rate_df(filter(clean_data, !str_detect(died_covid, "unknown|under review")), died_covid == "yes"))`, but importantly, `r fmt_count(clean_data, died_covid == "under review")` cases are under review, and `r fmt_count(clean_data, died_covid == "unknown")` of cases missing outcome status. 

## **Person** 

Since the start of the outbreak there was,  `r fmt_count(clean_data, case_gender == "female")` female  cases and `r fmt_count(clean_data, case_gender == "male")` male cases. 

The Age of cases ranges from `r min(clean_data$case_age, na.rm=T) ` to `r max(clean_data$case_age, na.rm=T)`, with Age group `r slice_tail(arrange(tabyl(clean_data$age_cat),n))[1]` the most represented, with 
``r round(digits =5, slice_tail(arrange(tabyl(clean_data$age_cat),n))[1,2])`` (`r round(digits = 2, slice_tail(arrange(tabyl(clean_data$age_cat),n))[1,3])`%) cases. 

There is `r fmt_count(clean_data, case_gender == "unknown")` cases with missing data on Sex, and `r fmt_count(clean_data, age_cat == "unknown")` cases with missing data on Age. 

### Age and Sex distributions

```{r Sex vs age group distribution }

clean_data %>% 
  tab_linelist(age_cat, strata = case_gender) %>% 
  
  select(-variable) %>% 
  rename("Age group" = value) %>% 
  rename_redundant("%" = proportion) %>% 
  augment_redundant("cases (n)" = "n$") %>% 
  
  #style kable
  kable(caption = "Distribution of cases per <b>Age groups</b> and <b>case gender</b>",
        align = "c",
        digits = 2) %>%
  
  kable_styling("hover", 
                full_width = F) 
  

```


#### Age-Sex pyramid

```{r}

```

### Hospitalisations 

This section investigates the Hospitalisations due to COVID throughout Age groups and sex. 

Overall Hospitalisation rates is of `r fmt_ci_df(case_fatality_rate_df(clean_data, hospitalized == "hospitalised"))`, but `r fmt_count(clean_data, hospitalized == "unknown")` still have an **unknown** status. 

#### Age group 

The age group, `r slice_tail(arrange(case_fatality_rate_df(clean_data, hospitalized == "hospitalised", group = age_cat), cfr))$age_cat` had the greatest hospitalisation rate of `r round(digits = 2, slice_tail(arrange(case_fatality_rate_df(clean_data, hospitalized == "hospitalised", group = age_cat), cfr))$cfr)`% `r slice_tail(arrange(case_fatality_rate_df(clean_data, hospitalized == "hospitalised", group = age_cat, mergeCI=T), cfr))$ci`. 

```{r hospitalisation rates per age_group }
#using the case_fatality_rate_df() function to calculate the Hospitalisation rates 

case_fatality_rate_df(clean_data, 
                      hospitalized == "hospitalised", 
                      group = age_cat, 
                      mergeCI = T) %>% 
  
  rename("hospitalisations" = deaths, 
         "Hosp. rate" = cfr, 
         "95%-CI" = ci, 
         "Age groups" = age_cat) %>% 
 
   kable(caption = "Hospitalisation rates per <b>Age groups</b>",
        align = "c",
        digits = 2) %>%
  
  kable_styling("hover", 
                full_width = F) 


```

#### Sex

Hopsitalisations rates remained similar between Male  (`r round(digits = 2, filter(case_fatality_rate_df(clean_data, hospitalized == "hospitalised", group = case_gender, mergeCI = T), case_gender == "male")$cfr)`% `r filter(case_fatality_rate_df(clean_data, hospitalized == "hospitalised", group = case_gender, mergeCI = T), case_gender == "male")$ci`) and Female (`r round(digits = 2, filter(case_fatality_rate_df(clean_data, hospitalized == "hospitalised", group = case_gender, mergeCI = T), case_gender == "female")$cfr)`% `r filter(case_fatality_rate_df(clean_data, hospitalized == "hospitalised", group = case_gender, mergeCI = T), case_gender == "female")$ci` )

```{r hospitalisation rates per Sex}
case_fatality_rate_df(clean_data, 
                      hospitalized == "hospitalised", 
                      group = case_gender, 
                      mergeCI = T) %>% 
  
  rename("hospitalisations" = deaths, 
         "Hosp. rate" = cfr, 
         "95%-CI" = ci, 
         "Sex" = case_gender) %>% 
 
   kable(caption = "Hospitalisation rates per <b> Sex </b>",
        align = "c",
        digits = 2) %>%
  
  kable_styling("hover", 
                full_width = F) 

```

### COVID Deaths 

This section looks into the CFR within Age groups and sex. 

Using only cases with known outcome, the Overall CFR is of `r fmt_ci_df(case_fatality_rate_df(filter(clean_data, !str_detect(died_covid, "unknown|under review")), died_covid == "yes"))`. 

Importantly, `r fmt_count(clean_data, died_covid == "under review")` cases are **under review**, while `r fmt_count(clean_data, died_covid == "unknown")` have unknown **outcome**. 

#### Age group 
```{r CFR per age group}
clean_data %>% 
  
    #filter out all cases with no outcome info
  filter(!str_detect(died_covid, "unknown|under review")) %>% 
  #calculate CFR per each group
  case_fatality_rate_df(
                        died_covid == "yes",
                        group= age_cat,
                        mergeCI = T) %>% 
  #rename
  rename("Age groups" = age_cat, 
         "CFR" = cfr, 
         "95%-CI" = ci) %>% 
  
  #style kable
  kable(caption = "Distribution of <b>COVID CFR</b> per <b>Age group</b>",
        align = "c",
        digits = 2) %>%
  
  kable_styling("hover", 
                full_width = F) 

```

#### Sex 

```{r CFR per sex }
clean_data %>% 
  #filter out all cases with no outcome info
  filter(!str_detect(died_covid, "unknown|under review")) %>% 
  
  #calculate CFR per group
  case_fatality_rate_df(
                        died_covid == "yes",
                        group= case_gender,
                        mergeCI = T) %>% 
  #rename vars
  rename("Sex" = case_gender, 
         "CFR" = cfr, 
         "95%-CI" = ci) %>% 
  
  #style kable
  kable(caption = "Distribution of <b>COVID CFR</b> per <b>Sex</b>",
        align = "c",
        digits = 2) %>%
  
  kable_styling("hover", 
                full_width = F) 
```

## **Time**
```{r date handling, eval = T, echo = F, include = F, warning = F, message = F}

# check the symptoms onset dates - many dates are missed typed 
clean_data %>% group_by(floor_date(sym_startdt_FALSE, "year")) %>% count()

#clean the symptoms onset dates 
#make NA all dates that are prior to Nov 2019 & make NA all date that are past the current date
clean_data <-  clean_data %>%
  mutate(sym_startdt_clean =  dplyr::if_else(sym_startdt_FALSE > as.Date("2019-11-01"), 
                                             
                                             if_else(sym_startdt_FALSE > Sys.Date(), 
                                                     as.Date(NA), 
                                                     sym_startdt_FALSE),
                                             
                                             as.Date(NA))) %>% 
  
  relocate(sym_startdt_clean, .after = sym_startdt_FALSE)

#check the confirmation dates - 122 missing dates but no out of boundaries
clean_data %>% group_by(floor_date(pos_sampledt_FALSE, "year")) %>% count()

```

### Main Epicurve
The peak of the outbreak was reached on week  ``r date2week(find_peak(inci), floor_day = T)``, with a total of ``r max(inci$count) `` cases.  ``r nrow(filter(clean_data, is.na(pos_sampledt_FALSE)))`` cases had missing dates, and they were excluded from the epicurves.

<center>
```{r main epicurve, echo = F}
#this is the main Epicurve - I need to display the accurate week names and pivot week labels
clean_data %>%
  
   group_by(weeks = ISOweek::ISOweek(floor_date(pos_sampledt_FALSE, "week", 
                                                    week_start = getOption("lubridate.week.start",1))))  %>% 

  
  count() %>%   
  
  #remove cases with no dates - 
  drop_na(weeks) %>% 
  
  hchart("column", hcaes(x = weeks, y = 'n')) %>% 
  
  hc_colors("#F4D497") %>% 
  
  hc_plotOptions(column = list(cropThreshold = 1000), 
                 series = list(groupPadding = 0, 
                               pointPadding = -0)) %>% 
  
  
  hc_xAxis(categories = unique(.$weeks)) %>% 
  
  hc_title(text = paste0("Cases by weeks of confirmation"), 
           align = "left", 
           style = list(useHTML =T, color = "#34495E", fontSize= 13)) %>%
  
  hc_tooltip(useHTML = T,
             formatter = JS("function(){
     outHTML =  '<b>' + this.point.weeks + '</b>' + '<br> <b> N cases: </b>' + this.point.n + '<br> <b>'
     return(outHTML)
     }"))

```
</center>

### Epicurves by groups
```{r epicurve function, include = F}

#define a function to create an interactive Epicurve where you can specify a grouping variable

epicurve <- function(clean_data, grouping_var, var_title) {
  
  #compute sum for each weeks
  week_sum <- clean_data %>%
  
   group_by(weeks = floor_date(pos_sampledt_FALSE, "week", 
                               week_start = getOption("lubridate.week.start",1))) %>% 
  
  count(name = "N") %>% mutate(isoweeks = ISOweek::ISOweek(weeks)) %>% relocate(isoweeks, .after = weeks)

clean_data %>%
  
   group_by(weeks = floor_date(pos_sampledt_FALSE, "week", 
                               week_start = getOption("lubridate.week.start",1)), 
            grouping_var) %>% 
  
  count(name = "n") %>% 
  
  left_join(., week_sum, by = "weeks") %>% 
  
  mutate(percent = round(digits = 2, n/N*100)) %>% 
  
  drop_na(weeks) %>% 
  
  hchart("column", hcaes(x = weeks, y = 'n', group = grouping_var), stacking = "normal") %>% 
  
  hc_plotOptions(column = list(cropThreshold = 1000), 
                 series = list(groupPadding = 0, 
                               pointPadding = -0)) %>% 
  
  
  hc_xAxis(categories = unique(.$weeks)) %>% 
  
  hc_title(text = paste0("Cases by week of confirmation and <b>", var_title), 
           align = "left", 
           style = list(useHTML =T, color = "#34495E", fontSize= 13)) %>%
    
    hc_tooltip(useHTML = T,
             formatter = JS(
               "
     function(){
     outHTML =  '<b>' + this.point.isoweeks + '</b>' + '<br><b>' + this.point.grouping_var + '<br>N cases: </b>' + this.point.n + '<br> <b> % of weekly cases: </b>' +this.point.percent 
     return(outHTML)
     }
     "))
  
}

```

<center>
```{r groups epicurves, echo = F}

hw_grid( 
  
  #Sex
  epicurve(clean_data = clean_data, grouping_var = clean_data$case_gender, var_title= "Sex") %>%
  
  hc_colors(c("#D6F2FA", "#EAD6FA", "#F9EDC2")), 
  
  #Age group
   epicurve(clean_data = clean_data, grouping_var = clean_data$age_cat, var_title= "Age categories") %>%
  
  hc_colors(c("#D5F5E3", "#ABEBC6", "#82E0AA", "#58D68D", "#2ECC71", "#239B56", "#1D8348", "#186A3B", "#17202A")), 
   
   #Diagnostic status - only one level but could be useful if suspected/probable cases added
    epicurve(clean_data = clean_data, grouping_var = clean_data$covid_dx, var_title= "diagnostic status") %>%
  
    hc_colors(c("#D6F2FA", "#EAD6FA", "#F9EDC2")), 
    
  #Hospitalisations
    epicurve(clean_data = clean_data, grouping_var = clean_data$hospitalized, var_title= "Hospitalisation status") %>%
  
    hc_colors(c("#A9DFBF", "#F5B7B1", "#F9EDC2" )) )
```
</center>

### Weekly Attack Rates, Hospitalisation rates and CFR


```{r}








```







## **Place** 
```{r spatial reference of clean_data, include = F}
#make clean_data spatial referenced - drop NA in long and lat var - set crs as WGS 1984
clean_data_sf <- clean_data %>% 
  
  drop_na(longitude_JITT, latitude_JITT) %>% 

  st_as_sf(coords = c("longitude_JITT", "latitude_JITT"), crs = 4326) 

#some cases have long and lat around 0 - probably typos need to be dealt with
hist(clean_data$longitude_JITT)
hist(clean_data$latitude_JITT)

#only keep cases which are found within the vector dataset - perform a spatial filter
clean_data_sf <- st_filter(clean_data_sf, vector)

```

### Cases map 

Cases have been detected throughout Fulton county, especially in major cities. The South-West part of the County

<center>
```{r cases map, message = F, echo = F}

#display map of cases to identify potentially undersampled locations - not the case here 

ggplot() +
  
  geom_sf(data = vector, fill = "#ffffff", col = "black", size = .2) +
  geom_sf(data = clean_data_sf, col = "#F8D18D", size = .1, alpha = .2) +
  
  ggtitle("Map of detected cases within Fulton county")+
  
  theme(plot.margin = unit(c(2, 2, 2, 2), "mm"), 
        panel.background = element_blank(), 
        legend.background=element_blank(),
        axis.line = element_blank(), 
        axis.title = element_blank(), 
        axis.ticks = element_blank(), 
        axis.text = element_blank(), 
        plot.title = element_text(size = 13, 
                                  color = "#34495E", 
                                  hjust = 0, 
                                  vjust = 0)) +
  ggsn::scalebar(data=vector, dist= 15, dist_unit = "km", 
           transform=T, 
           model="WGS84", 
           border.size = .2, 
           st.size = 2, 
           height = .01, 
           location= "bottomleft", 
           anchor = c(x=-84.4, y =33.5 )
           ) + 
    
    ggsn::north(data=vector, 
          symbol = 14, 
          scale= .08, 
          location= "topleft")  
  
 

```
</center>

### Chloropleths maps 

Below are displayed the Attack rates within Zip codes of Fulton country. 

```{r data aggregating}


subset_clean_data <- clean_data %>% dplyr::select(case_zip, died_covid, hospitalized, pos_sampledt_FALSE)

#extract population data to perform aggregation - remove geometry for now
pop_data <- vector %>% 
  dplyr::select(ZipCode, Population) %>% 
  st_drop_geometry()

#join pop data and epi data 

subset_clean_data <- left_join(subset_clean_data, pop_data, by = c("case_zip" = "ZipCode"))

subset_clean_data %>% group_by(weeks = floor_date(pos_sampledt_FALSE, "week", 
                               week_start = getOption("lubridate.week.start",1)), 
                               case_zip) %>%
  
  summarise(AR = n()/Population * 10000) 
  
#join spatial data and epi data 




```
