---
title: "Epidemiological Report - COVID-19"
author: "Hugo Soubrier"
date: "``r Sys.Date()``"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
---

```{r set up, include = F}

knitr::opts_chunk$set(echo = FALSE,       
                      error = TRUE,
                      message = FALSE, 
                      warning = F)

```

```{r library loading}
#spatial analysis 
library(shiny)
library(sf)
library(ggsn)
library(raster)
library(leaflet)

#epidemiological analysis 

library(ISOweek)
library(shinyWidgets)
library(sitrep)
library(skimr)
library(epikit)
library(incidence)
library(linelist)
library(magrittr)
library(utils)
library(lubridate)
library(aweek)
library(naniar)
library(ggplot2)
library(devtools)
library(plotly)
library(shinydashboard)
library(shinyjs)
library(janitor)
library(lubridate)
library(sunburstR)
library(knitr)
library(kableExtra)
library(rio)
library(here)
library(tidyselect)
library(tidyverse)
library(highcharter)
```
## Introduction 

```{r data loading & cleaning,}
#read data
data <- import(here("data", "covid_example_data.xlsx"))

# set data classes -----------------------------------

#create vector for the factors variables
fct_vars <- data %>% select(contains("sym"), #select all symptoms var
                            -sym_startdt_FALSE, #exclude the symptoms date var 
                            - sym_resolveddt_FALSE, #exclude the symptoms resolved date var 
                            case_gender, 
                            case_race, 
                            case_eth, 
                            hospitalized, 
                            died, 
                            died_covid, 
                            confirmed_case, 
                            covid_dx) %>% names()

#create vector for the date variables
date_vars <- data %>% select(case_dob_FALSE, contains("dt")) %>% names()

#create vector for numeric variables
numeric_vars <- data %>% select(case_zip, case_age, latitude_JITT, longitude_JITT) %>% names()

#create vector for character variables
character_vars <- data %>% select(PID) %>% names()


#apply data class to each vectors -
clean_data <- data %>% mutate(across(fct_vars, as_factor),
                across(date_vars, as_date), 
                across(numeric_vars, as.numeric), 
                across(character_vars, as.character))

# Tidy data -------------------
#create vector of var that need cleaning 
needtidy_vars <- clean_data %>% select(where(is.factor)) %>% colnames()

# use linelist::clean_data on it 
clean_data[needtidy_vars] <- clean_data %>% select(all_of(needtidy_vars)) %>% clean_data()

#Age 
#make NA all obs that have an age < 0
clean_data <- clean_data %>% mutate(case_age = ifelse(case_age<0, NA, case_age))

```

```{r spatial data loading, include = F}
#read spatial data and make the ZipCode variable the same type as the case_zip in the Epi data for joining - numeric
vector <- read_sf("data/vectors/FultonCountyZipCodes.shp") %>% 
  mutate(ZipCode = as.numeric(ZipCode))

#set the CRS to 4326 - WGS 1984
vector <- st_transform(vector, 4326)

#extract population data to perform ARs - remove geometry for now
pop_data <- vector %>% 
  dplyr::select(ZipCode, Population) %>% 
  st_drop_geometry()

#County 30331 is duplicated in the shapefile - need to remove one otherwise will be dupicated on map 
pop_data %>% get_dupes()

pop_data <- pop_data %>% filter(!duplicated(ZipCode))

```

```{r data recoding & classifying, include = F}
# create age_cat var 
clean_data <- clean_data %>%  mutate(age_cat = case_when( case_age < 10 ~ "< 10", 
                                                between(case_age, 10, 20) ~ "10 - 20", 
                                                between(case_age, 20, 30) ~ "20 - 30", 
                                                between(case_age, 30, 40) ~ "30 - 40", 
                                                between(case_age, 40, 50) ~ "40 - 50", 
                                                between(case_age, 50, 60) ~ "50 - 60", 
                                                between(case_age, 60, 70) ~ "60 - 70", 
                                                case_age > 70 ~ "70 +"), 
                                     age_cat = if_else(is.na(age_cat), "unknown", age_cat), 
                                     age_cat = fct_relevel(age_cat, c("< 10", "10 - 20", "20 - 30", "30 - 40", "40 - 50", "50 - 60", "60 - 70", "70 +"))) %>% 
  
  relocate(age_cat, .after = case_age)

#recode all variables - if NA then set as unknown  

clean_data <- clean_data %>% mutate(across(fct_vars, as.character),
                      across(fct_vars, ~if_else(is.na(.x), "unknown", .x)),
                      across(fct_vars, as_factor))

#recode specific variables

clean_data <- clean_data %>% 
  mutate( hospitalized = recode(hospitalized, 
                                "yes" = "hospitalised",
                                "no" = "not hospitalised"))

clean_data <- clean_data %>% 
  mutate(died_covid = recode( died_covid, 
                              "under_review" = "under review"))


```

```{r incidence calculation, include = F}
#calculate incidence and get ISO-weeks for the outbreak for later uses
inci <- incidence(clean_data$pos_sampledt_FALSE, interval = "monday week")

```

## **Overview** 
Between the  `r min(inci$dates)` and the `r max(inci$dates)` (a period `r inci$timespan ` days) a total of `r inci$n` cases have been identified, of which `r fmt_count(clean_data, covid_dx == "confirmed")` are **confirmed** cases. 

- In Fulton County, and over this period, the **attack rate** is of `r fmt_ci_df(attack_rate(inci$n, population = sum(pop_data$Population), multiplier = 10000), percent = F)` per 10 000.

- Amongst all cases of known hospitalisation status, at least `r fmt_count(filter(clean_data, hospitalized !="unknown"), hospitalized == "hospitalised")` have been **hospitalised**, but `r fmt_count(clean_data, hospitalized == "unknown")` still have an **unknown** status 

- Using only cases with known outcome, the **Case Fatality Ratio** is of `r fmt_ci_df(case_fatality_rate_df(filter(clean_data, !str_detect(died_covid, "unknown|under review")), died_covid == "yes"))`, but importantly, `r fmt_count(clean_data, died_covid == "under review")` cases are under review, and `r fmt_count(clean_data, died_covid == "unknown")` of cases missing outcome status. 

## **Person** 

Since the start of the outbreak there was,  `r fmt_count(clean_data, case_gender == "female")` female  cases and `r fmt_count(clean_data, case_gender == "male")` male cases. 

The Age of cases ranges from `r min(clean_data$case_age, na.rm=T) ` to `r max(clean_data$case_age, na.rm=T)`, with Age group `r slice_tail(arrange(tabyl(clean_data$age_cat),n))[1]` the most represented, with 
``r round(digits =2, slice_tail(arrange(tabyl(clean_data$age_cat),n))[1,2])`` (`r round(digits = 2, slice_tail(arrange(tabyl(clean_data$age_cat),n))[1,3]*100)`%) cases. 

There is `r fmt_count(clean_data, case_gender == "unknown")` cases with missing data on Sex, and `r fmt_count(clean_data, age_cat == "unknown")` cases with missing data on Age. 

### Missing data



### Age and Sex distributions

```{r Sex vs age group distribution }

clean_data %>% 
  tab_linelist(age_cat, strata = case_gender) %>% 
  
  select(-variable) %>% 
  rename("Age group" = value) %>% 
  rename_redundant("%" = proportion) %>% 
  augment_redundant("cases (n)" = "n$") %>% 
  
  #style kable
  kable(caption = "Distribution of cases per <b>Age groups</b> and <b>case gender</b>",
        align = "c",
        digits = 2) %>%
  
  kable_styling("hover", 
                full_width = F) 
  

```


#### Age-Sex pyramid

```{r}

```

### Hospitalisations 

Overall Hospitalisation rates is of `r fmt_ci_df(case_fatality_rate_df(filter(clean_data, hospitalized !="unknown"), hospitalized == "hospitalised"))`, but `r fmt_count(clean_data, hospitalized == "unknown")` still have an **unknown** status. 

#### Age group 

The age group, `r slice_tail(arrange(case_fatality_rate_df(filter(clean_data, hospitalized !="unknown"), hospitalized == "hospitalised", group = age_cat), cfr))$age_cat` had the greatest hospitalisation rate of `r round(digits = 2, slice_tail(arrange(case_fatality_rate_df(filter(clean_data, hospitalized !="unknown"), hospitalized == "hospitalised", group = age_cat), cfr))$cfr)`% `r slice_tail(arrange(case_fatality_rate_df(filter(clean_data, hospitalized !="unknown"), hospitalized == "hospitalised", group = age_cat, mergeCI=T), cfr))$ci`. 

```{r hospitalisation rates per age_group }
#using the case_fatality_rate_df() function to calculate the Hospitalisation rates 

case_fatality_rate_df(filter(clean_data, hospitalized !="unknown"), 
                      hospitalized == "hospitalised", 
                      group = age_cat, 
                      mergeCI = T) %>% 
  
  rename("hospitalisations" = deaths, 
         "Hosp. rate" = cfr, 
         "95%-CI" = ci, 
         "Age groups" = age_cat) %>% 
 
   kable(caption = "Hospitalisation rates per <b>Age groups</b>",
        align = "c",
        digits = 2) %>%
  
  kable_styling("hover", 
                full_width = F) 


```

#### Sex

Hopsitalisations rates remained similar between **Male**  (`r round(digits = 2, filter(case_fatality_rate_df(filter(clean_data, hospitalized !="unknown"), hospitalized == "hospitalised", group = case_gender, mergeCI = T), case_gender == "male")$cfr)`% `r filter(case_fatality_rate_df(filter(clean_data, hospitalized !="unknown"), hospitalized == "hospitalised", group = case_gender, mergeCI = T), case_gender == "male")$ci`) and **Female** (`r round(digits = 2, filter(case_fatality_rate_df(filter(clean_data, hospitalized !="unknown"), hospitalized == "hospitalised", group = case_gender, mergeCI = T), case_gender == "female")$cfr)`% `r filter(case_fatality_rate_df(filter(clean_data, hospitalized !="unknown"), hospitalized == "hospitalised", group = case_gender, mergeCI = T), case_gender == "female")$ci` )

```{r hospitalisation rates per Sex}

case_fatality_rate_df(filter(clean_data, hospitalized !="unknown"), 
                      hospitalized == "hospitalised", 
                      group = case_gender, 
                      mergeCI = T) %>% 
  
  rename("hospitalisations" = deaths, 
         "Hosp. rate" = cfr, 
         "95%-CI" = ci, 
         "Sex" = case_gender) %>% 
 
   kable(caption = "Hospitalisation rates per <b> Sex </b>",
        align = "c",
        digits = 2) %>%
  
  kable_styling("hover", 
                full_width = F) 

```

### COVID Deaths 

Using only cases with known outcome, the Overall CFR is of `r fmt_ci_df(case_fatality_rate_df(filter(clean_data, !str_detect(died_covid, "unknown|under review")), died_covid == "yes"))`. 

Importantly, `r fmt_count(clean_data, died_covid == "under review")` cases are **under review**, while `r fmt_count(clean_data, died_covid == "unknown")` have unknown **outcome**. 

#### Age group 
```{r CFR per age group}
age_cat_cfr <- clean_data %>% 
  
    #filter out all cases with no outcome info
  filter(!str_detect(died_covid, "unknown|under review")) %>% 
  
  #calculate CFR per each group - do not merge CI here to be used later in inline function
  case_fatality_rate_df(died_covid == "yes",
                        group= age_cat,
                        mergeCI = F)
  #rename
age_cat_cfr %>% 
  
  #merge CI 
  mutate(lower = round(digits = 2, lower), 
         upper = round(digits = 2, upper),
    "95%-CI" = paste0("(", lower, "-", upper, ")")) %>% 
  
  select(-c(lower, upper)) %>% 
  
  
  rename("Age groups" = age_cat, 
         "CFR" = cfr) %>% 
  
  #style kable
  kable(caption = "Distribution of COVID <b>CFR</b> per <b>Age group</b>",
        align = "c",
        digits = 2) %>%
  
  kable_styling("hover", 
                full_width = F) 
```

**`r filter(filter(age_cat_cfr, age_cat !="unknown"), cfr == min(age_cat_cfr$cfr))$age_cat`** was the age group with the lowest CFR: `r fmt_ci_df(filter(filter(age_cat_cfr, age_cat !="unknown"), cfr == min(age_cat_cfr$cfr)), e = 4, l =5, u = 6)`. While **`r filter(filter(age_cat_cfr, age_cat !="unknown"), cfr == max(age_cat_cfr$cfr))$age_cat`** was the age group with the greatest CFR: `r fmt_ci_df(filter(filter(age_cat_cfr, age_cat !="unknown"), cfr == max(age_cat_cfr$cfr)), e = 4, l =5, u = 6) `

#### Sex 

```{r CFR per sex }

clean_data %>% 
  #filter out all cases with no outcome info
  filter(!str_detect(died_covid, "unknown|under review")) %>% 
  
  #calculate CFR per group
  case_fatality_rate_df(
                        died_covid == "yes",
                        group= case_gender,
                        mergeCI = T) %>% 
  #rename vars
  rename("Sex" = case_gender, 
         "CFR" = cfr, 
         "95%-CI" = ci) %>% 
  
  #style kable
  kable(caption = "Distribution of COVID <b>CFR</b> per <b>Sex</b>",
        align = "c",
        digits = 2) %>%
  
  kable_styling("hover", 
                full_width = F) 
```

## **Time**
```{r date handling, eval = T, echo = F, include = F, warning = F, message = F}

# check the symptoms onset dates - many dates are missed typed 
clean_data %>% group_by(floor_date(sym_startdt_FALSE, "year")) %>% count()

#clean the symptoms onset dates 
#make NA all dates that are prior to Nov 2019 & make NA all date that are past the current date
clean_data <-  clean_data %>%
  mutate(sym_startdt_clean =  dplyr::if_else(sym_startdt_FALSE > as.Date("2019-11-01"), 
                                             
                                             if_else(sym_startdt_FALSE > Sys.Date(), 
                                                     as.Date(NA), 
                                                     sym_startdt_FALSE),
                                             
                                             as.Date(NA))) %>% 
  
  relocate(sym_startdt_clean, .after = sym_startdt_FALSE)

#check the confirmation dates - 122 missing dates but no out of boundaries
clean_data %>% group_by(floor_date(pos_sampledt_FALSE, "year")) %>% count()
```

### Main Epicurve
The peak of the outbreak was reached on week  ``r date2week(find_peak(inci), floor_day = T)``, with a total of ``r max(inci$count) `` cases.  ``r nrow(filter(clean_data, is.na(pos_sampledt_FALSE)))`` cases had missing dates, and they were excluded from the epicurves.

<center>
```{r main epicurve, echo = F}
#this is the main Epicurve - I need to display the accurate week names and pivot week labels
clean_data %>%
  
   group_by(weeks = ISOweek::ISOweek(floor_date(pos_sampledt_FALSE, "week", 
                                                    week_start = getOption("lubridate.week.start",1))))  %>% 

  
  count() %>%   
  
  #remove cases with no dates - 
  drop_na(weeks) %>% 
  
  hchart("column", hcaes(x = weeks, y = 'n')) %>% 
  
  hc_colors("#F4D497") %>% 
  
  hc_plotOptions(column = list(cropThreshold = 1000), 
                 series = list(groupPadding = 0, 
                               pointPadding = -0)) %>% 
  
  
  hc_xAxis(categories = unique(.$weeks)) %>% 
  
  hc_title(text = paste0("Cases by weeks of confirmation"), 
           align = "left", 
           style = list(useHTML =T, color = "#34495E", fontSize= 13)) %>%
  
  hc_tooltip(useHTML = T,
             formatter = JS("function(){
     outHTML =  '<b>' + this.point.weeks + '</b>' + '<br> <b> N cases: </b>' + this.point.n + '<br> <b>'
     return(outHTML)
     }")) %>% 
  
  hc_navigator( enabled = TRUE, 
                xAxis = list(
                  labels = list(
                    enabled = F
                  )
                )
                )

```
</center>

### Epicurves by groups
```{r epicurve function, include = F}

#define a function to create an interactive Epicurve where you can specify a grouping variable

epicurve <- function(clean_data, grouping_var, var_title) {
  
  #compute sum for each weeks
  week_sum <- clean_data %>%
  
   group_by(weeks = floor_date(pos_sampledt_FALSE, "week", 
                               week_start = getOption("lubridate.week.start",1))) %>% 
  
  count(name = "N") %>% mutate(isoweeks = ISOweek::ISOweek(weeks)) %>% relocate(isoweeks, .after = weeks)

clean_data %>%
  
   group_by(weeks = floor_date(pos_sampledt_FALSE, "week", 
                               week_start = getOption("lubridate.week.start",1)), 
            grouping_var) %>% 
  
  count(name = "n") %>% 
  
  left_join(., week_sum, by = "weeks") %>% 
  
  mutate(percent = round(digits = 2, n/N*100)) %>% 
  
  drop_na(weeks) %>% 
  
  hchart("column", hcaes(x = weeks, y = 'n', group = grouping_var), stacking = "normal") %>% 
  
  hc_plotOptions(column = list(cropThreshold = 1000), 
                 series = list(groupPadding = 0, 
                               pointPadding = -0)) %>% 
  
  
  hc_xAxis(categories = unique(.$weeks)) %>% 
  
  hc_title(text = paste0("Cases by week of confirmation and <b>", var_title), 
           align = "left", 
           style = list(useHTML =T, color = "#34495E", fontSize= 13)) %>%
    
    hc_tooltip(useHTML = T,
             formatter = JS(
               "
     function(){
     outHTML =  '<b>' + this.point.isoweeks + '</b>' + '<br><b>' + this.point.grouping_var + '<br>N cases: </b>' + this.point.n + '<br> <b> % of weekly cases: </b>' +this.point.percent 
     return(outHTML)
     }
     "))
  
}

```

<center>
```{r groups epicurves, echo = F}

hw_grid( 
  
  #Sex
  epicurve(clean_data = clean_data, grouping_var = clean_data$case_gender, var_title= "Sex") %>%
  
  hc_colors(c("#D6F2FA", "#EAD6FA", "#F9EDC2")), 
  
  #Age group
   epicurve(clean_data = clean_data, grouping_var = clean_data$age_cat, var_title= "Age categories") %>%
  
  hc_colors(c("#D5F5E3", "#ABEBC6", "#82E0AA", "#58D68D", "#2ECC71", "#239B56", "#1D8348", "#186A3B", "#17202A")), 
   
   #Diagnostic status - only one level but could be useful if suspected/probable cases added
    epicurve(clean_data = clean_data, grouping_var = clean_data$covid_dx, var_title= "diagnostic status") %>%
  
    hc_colors(c("#D6F2FA", "#EAD6FA", "#F9EDC2")), 
    
  #Hospitalisations
    epicurve(clean_data = clean_data, grouping_var = clean_data$hospitalized, var_title= "Hospitalisation status") %>%
  
    hc_colors(c("#A9DFBF", "#F5B7B1", "#F9EDC2" )) )
```
</center>

### Weekly Attack Rates, Hospitalisation rates and CFR

```{r compute weekly attack rate, Hosp. rate and CFR}

#calculate N cases per weeks, required for attack_rate()
weekly_cases <- clean_data %>% 
  
  drop_na(pos_sampledt_FALSE) %>% #remove samples that do not have a valid date of confirmation
  
  group_by(ISOweek = ISOweek::ISOweek(floor_date(pos_sampledt_FALSE, "weeks")))  %>% 
  
  count() %>% 
  
  #add pop data of Fulton county
  mutate(population = sum(pop_data$Population))

#AR
weekly_ar <- attack_rate(weekly_cases$n, 
              population = weekly_cases$population, 
              mergeCI= F, 
              multiplier = 10000) %>% 
    
    bind_cols(select(weekly_cases, ISOweek)) %>% #add ISOweek variable
    #set as first var
    relocate(ISOweek, 1) %>% 
  
  #change digits number for estimates and CI 
  mutate(across(c(ar, lower, upper), ~round(.x, digits = 2)))
  


#weekly Hosp. rate - calculated using the case_fatality_rate()
weekly_hosp_rate <- clean_data %>% 
  
  arrange(pos_sampledt_FALSE) %>% #arrange dates from earlier - latest
  
  mutate(ISOweek = as_factor(ISOweek(pos_sampledt_FALSE)))  %>% #create an ISOweek var for graph
  
  filter(hospitalized != "unknown") %>% #remove all not hospitalised

case_fatality_rate_df(., #compute hospitalisation rates grouped by epiweeks 
                      hospitalized == "hospitalised", 
                      group = ISOweek, 
                      mergeCI = F) %>% 
  #rename var accordingly
  rename("hospitalisations" = deaths, 
         "hosp_rate" = cfr) %>% 
  
  #change digits number for estimates and CI 
  mutate(across(c(hosp_rate, lower, upper), ~round(.x, digits = 2)))




#CFR - same logic as hospitalisation rates
weekly_cfr <- clean_data %>% 
  
  arrange(pos_sampledt_FALSE) %>% 
  
  mutate(ISOweek = as_factor(ISOweek(pos_sampledt_FALSE)))  %>% 
  
  filter(!str_detect(died_covid, "unknown|under review")) %>% #remove all with unknown/under review 
  
  arrange(ISOweek) %>% 

case_fatality_rate_df(., 
                      died_covid == "yes", 
                      group = ISOweek, 
                      mergeCI = F) %>%
  
  #change digits number for estimates and CI 
  mutate(across(c(cfr, lower, upper), ~round(.x, digits = 2)))

```


```{r AR, CFR, and hosp rate tables, eval = F}
#weekly Tables for rates -- not evaluated as all info are on the plot 

#weekly AR table
 weekly_ar %>%  
   
   merge_ci_df(e = 4,
               l = 5, 
               u =6) %>% 
   rename( "95%-CI" = ci, 
           "AR per 10000" = ar ) %>% 
   
   kable(caption = "Attack Rates per <b>ISOweeks</b> of the outbreak",
        align = "c",
        digits = 3) %>%
  
  kable_styling("hover", 
                full_width = F)


#weekly Hospi. rate table 
weekly_hosp_rate  %>%  
  
  merge_ci_df(e =4, 
              l =5, 
              u =6) %>% 
  
  rename("95%-CI" = ci, 
         "Hosp. rate" = hosp_rate) %>% 
  
  
  kable(caption = "Hospitalisation Rates per <b>ISOweeks</b> of the outbreak",
        align = "c",
        digits = 3) %>%
  
  kable_styling("hover", 
                full_width = F)

#weekly CFR table 
weekly_cfr %>%  
  
  merge_ci_df(e =4, 
              l =5, 
              u =6 ) %>% 
  
  rename("95%-CI" = ci, 
         "CFR" = cfr) %>% 
  
  kable(caption = "Case Fatality Rates per <b>ISOweeks</b> of the outbreak",
        align = "c",
        digits = 3) %>%
  
  kable_styling("hover", 
                full_width = F)






```

<center>
```{r rates graph}
highchart() %>% 
  
  #plot the AR 
  hc_add_series(name = "Attack_rate", 
                data = weekly_ar, 
                "line", 
                hcaes(x = ISOweek, 
                      y = ar), 
                color = "#5496D9", 
                fillOpacity= .8, 
                marker = list(enabled = F, 
                              radius = 2, 
                              symbol = "circle"), 
                linkedT0 = "AR 95%-CI") %>% 
  #AR CI
  hc_add_series(name = "AR 95%-CI",
                type = "arearange",
                data = weekly_ar,
                hcaes(x = ISOweek, 
                      low = weekly_ar$lower, 
                      high = weekly_ar$upper ), 
                linkedTo = "Attack rate", 
                color = "#A9CDF1", 
                fillOpacity= .2, 
                marker = list(enabled = F, 
                              radius = 2, 
                              symbol = "circle"), 
                lineWidth = 0) %>% 
  
  #Hosp.  rate
  hc_add_series(name = "Hospitalisations rate", 
                data = weekly_hosp_rate, 
                "line", 
                hcaes(x = ISOweek, 
                      y = hosp_rate), 
                color = "#EAD03C", 
                fillOpacity= .8, 
                marker = list(enabled = F, 
                              radius = 2, 
                              symbol = "circle")) %>% 
  #Hosp rate - CI
  hc_add_series(name = "Hosp.rate 95%-CI",
                type = "arearange",
                data = weekly_hosp_rate,
                hcaes(x = ISOweek, 
                      low = weekly_hosp_rate$lower, 
                      high = weekly_hosp_rate$upper ), 
                linkedTo = "Hospitalisation rate", 
                color = "#DCD298", 
                fillOpacity= .2, 
                lineWidth = 0, 
                marker = list(enabled = F, 
                              radius = 2, 
                              symbol = "circle")) %>% 
  #CFR
  hc_add_series(name = "CFR", 
                data = weekly_cfr, 
                "line", 
                hcaes(x = ISOweek, 
                      y = cfr), 
                color = "#E84F4F", 
                fillOpacity= .8, 
                marker = list(enabled = F,
                              radius = 2, 
                              symbol = "circle")) %>% 
  #CFR - CI
  hc_add_series(name = "CFR 95%-CI",
                type = "arearange",
                data = weekly_cfr,
                hcaes(x = ISOweek, 
                      low = weekly_cfr$lower, 
                      high = weekly_cfr$upper ), 
                linkedTo = "CFR", 
                lineWidth = 0,
                color = "#F8C8C8", 
                fillOpacity= .2, 
                zIndex = 0, 
                marker = list(enabled = F, 
                              radius = 2, 
                              symbol = "circle")) %>%
  
  hc_xAxis(categories = unique(weekly_ar$ISOweek)) %>% 
  
 hc_title(text = paste0("Attack Rate, Case Fatality Ratio and Hospitalisation rate by weeks of confirmation"), 
           align = "left", 
           style = list(useHTML =T, color = "#34495E", fontSize= 13)) %>% 
  
  hc_tooltip(shared = T,
             crosshairs = T,
             useHTML = T
             )
```
</center>



## **Place** 
```{r spatial reference of clean_data, include = F}
#make clean_data spatial referenced - drop NA in long and lat var - set crs as WGS 1984
clean_data_sf <- clean_data %>% 
  
  drop_na(longitude_JITT, latitude_JITT) %>% 

  st_as_sf(coords = c("longitude_JITT", "latitude_JITT"), crs = 4326) 

#some cases have long and lat around 0 - probably typos need to be dealt with
hist(clean_data$longitude_JITT)
hist(clean_data$latitude_JITT)

#only keep cases which are found within the vector dataset - perform a spatial filter
clean_data_sf <- st_filter(clean_data_sf, vector)

```

### Cases map 

Cases have been detected throughout Fulton county, especially in major cities. The South-West part of the County

<center>
```{r cases map, message = F, echo = F}

#display map of cases to identify potentially undersampled locations - not the case here 

ggplot() +
  
  geom_sf(data = vector, fill = "#ffffff", col = "black", size = .2) +
  geom_sf(data = clean_data_sf, col = "#F8D18D", size = .1, alpha = .2) +
  
  ggtitle("Map of detected cases within Fulton county")+
  
  theme(plot.margin = unit(c(2, 2, 2, 2), "mm"), 
        panel.background = element_blank(), 
        legend.background=element_blank(),
        axis.line = element_blank(), 
        axis.title = element_blank(), 
        axis.ticks = element_blank(), 
        axis.text = element_blank(), 
        plot.title = element_text(size = 13, 
                                  color = "#34495E", 
                                  hjust = 0, 
                                  vjust = 0)) +
  ggsn::scalebar(data=vector, dist= 15, dist_unit = "km", 
           transform=T, 
           model="WGS84", 
           border.size = .2, 
           st.size = 2, 
           height = .01, 
           location= "bottomleft", 
           anchor = c(x=-84.4, y =33.5 )
           ) + 
    
    ggsn::north(data=vector, 
          symbol = 14, 
          scale= .08, 
          location= "topleft")  
  
 

```
</center>

### Chloropleths maps 
```{r AR per week and ZIP, include = F}

#using clean_data, 1. calculate N cases per week and County, 2. We calculate the Attack rates per week and county 3. We subset AR tables in small and big regions because of the outliers. 4. We categorise the AR for mapping, 4. We spatial reference the AR tables using zip codes for mapping

#1. Calculate N cases per weeks and ZIP codes, required for attack_rate() - join the pop of each county

weekly_zip_cases <- clean_data %>% 
  
  filter(case_zip %in% pop_data$ZipCode) %>% #only keep cases that are included in the shapefile
  
  drop_na(pos_sampledt_FALSE) %>% #remove samples that do not have a valid date of confirmation
  
  group_by(ISOweek = ISOweek::ISOweek(floor_date(pos_sampledt_FALSE, "weeks")), 
           #grouping by ISOweek but also by month to later facet maps on months
           month = month(floor_date(pos_sampledt_FALSE, "month"), label = T), case_zip)  %>% 
  
  count() %>%
  
  #format the month var for later
  mutate(month = paste0(str_remove(ISOweek, "-W.*"), "-", month)) %>% 
  
  ungroup() %>% 

  #add pop of each ZIP
left_join(., pop_data, by= c("case_zip" = "ZipCode")) %>% 

#remove Zip code where pop is <0
filter(Population !=0)

vector %>% filter(ZipCode == 30349)

weekly_zip_cases %>% filter(case_zip == 30349)

# -----------------------------------------------

#2. Calculate AR for each week and Zip
weekly_zip_ar <- attack_rate(cases = weekly_zip_cases$n, 
            population = weekly_zip_cases$Population, 
            multiplier = 10000, 
            mergeCI = T) %>% 
  
  bind_cols(select(weekly_zip_cases, ISOweek, case_zip, month)) %>% #add ISOweek - Zip variable -  month
  
    #set as first var
    relocate(c(ISOweek, month, case_zip), 1) 

# -----------------------------------------------------------------------
#3. subset small regions 

#see distribution of AR - massive outliers that dilute visualisations (AR > 300) - located in  THREE regions with SMALL populations
hist(weekly_zip_ar$ar, breaks = 1000)
filter(weekly_zip_ar, ar > 300) %>% count(case_zip, population)

#create a subset using population size <1000
weekly_zip_ar_small <- weekly_zip_ar %>% filter(population<1000)

#remove them from the main dataset 
weekly_zip_ar_big <- weekly_zip_ar %>% filter(!(case_zip %in% weekly_zip_ar_small$case_zip ))


#----------------------------------------------------------------------
#4. Categorisation of ARs

## creates breaks for categories - main breaks and then breaks for small regions (outliers) -- not sure this is ideal but trying to get best nuances for visualisation

breaks <- as.integer(c(0, find_breaks(max(weekly_zip_ar_big$ar, na.rm=T), breaks = 5, snap = 5)))

breaks_small <- as.integer(c(0, find_breaks(max(weekly_zip_ar_small$ar, na.rm=T), breaks = 5, snap = 10)))

#categorize AR - main
weekly_zip_ar_big <- weekly_zip_ar_big %>% mutate(ar_cat = age_categories(ar, breaks))

#categorize AR - small regions
weekly_zip_ar_small <- weekly_zip_ar_small %>% mutate(ar_cat = age_categories(ar, breaks_small))

# -----------------------------------------------------------------------
#5. Geo-referencing the AR table using the geometry of shapefile - in vector

#add geometry to AR per zip codes 
weekly_zip_ar_sf_big <- left_join(weekly_zip_ar_big, select(vector, ZipCode, geometry), by = c("case_zip" = "ZipCode")) %>% 
  
  st_as_sf()

weekly_zip_ar_sf_small <- left_join(weekly_zip_ar_small, select(vector, ZipCode, geometry), by = c("case_zip" = "ZipCode")) %>% 
  
  st_as_sf()
```
 
#### Peak of Outbreak
<center>
```{r chloropleth peak}
#get peak of outbreak
peak <- date2week(find_peak(inci), floor_day = T)

#define palette 
pal <- colorFactor(palette = c("#94d2bd", 
                               "#ffe169", 
                               "#edc531", 
                               "#e85d04", 
                               "#dc2f02", 
                               "#bb3e03"), 
                    level = weekly_zip_ar_big$ar_cat, 
                    ordered = T)



#create Labels in HTML for hovering on map 
weekly_zip_ar_sf_big$labels <- paste0(weekly_zip_ar_sf_big$ISOweek, "<br>ZIP: <b>", weekly_zip_ar_sf_big$case_zip, "<br>", "Attack Rate: </b>", round(digits = 2, weekly_zip_ar_sf_big$ar), " per 10 000<br> <b>Population:</b> ", weekly_zip_ar_sf_big$population) %>% lapply(htmltools::HTML)


# map using leaflet 
leaflet(data = filter(filter(weekly_zip_ar_sf_big), ISOweek == peak), 
        
        options = leafletOptions(dragging =F, 
                                 minZoom = 9.5, 
                                 maxZoom = 9.5, 
                                 zoomControl = F)) %>% 
  
  addPolygons(fillColor = ~pal(ar_cat),
              weight  = 0.4,
              color = "black", 
              highlightOptions = highlightOptions(weight = 4, 
                                                  color = "red", 
                                                  bringToFront = T, 
                                                  opacity = .5), 
              label = ~labels) %>% 
  
  addLegend(pal = pal, values = ~ar_cat, 
            title = "Attack Rate<br><span style='font-size:10px;'>per 10 000") %>% 
  
  addControl(paste0(html = "<span style ='font-size:12px;color:#34495E;'>Chloropleth map of Fulton county at peak of the outbreak: <b>", peak, "</b></span><br><span style ='font-size:11px;color:#34495E;'>regions with more than 1000 habitants"))

```
</center>

<hr>

Below are mapped the smaller regions of Fulton county (<1000 habitants) which AR are really important (mean: `r round(digits = 2, mean(weekly_zip_ar_sf_small$ar))`) due to small size populations, mean`r round(digits = 2, mean(weekly_zip_ar_sf_small$population))`. This was done to allow a more nuanced mapping of other bigger regions by removing outliers. 

<hr>

<center>
```{r chloropleth peak small regions}

#I do this graph otherwise these regions make the distribution of AR extremely righ skewed and visaulisation is really hard on main graph due to massive outliers

#define palette 
pal_small <- colorFactor(palette = c( 
                               "#edc531", 
                               "#e85d04", 
                               "#dc2f02", 
                               "#bb3e03"), 
                    level = weekly_zip_ar_sf_small$ar_cat, 
                    ordered = T)

#create Labels in HTML for hovering on map 
weekly_zip_ar_sf_small$labels <- paste0(weekly_zip_ar_sf_small$ISOweek, "<br>ZIP: <b>", weekly_zip_ar_sf_small$case_zip, "<br>", "Attack Rate: </b>", round(digits = 2, weekly_zip_ar_sf_small$ar), " per 10 000<br> <b>Population:</b> ", weekly_zip_ar_sf_small$population) %>% lapply(htmltools::HTML)


# map using leaflet 
leaflet(data = filter(filter(weekly_zip_ar_sf_small), ISOweek == peak), 
        
        options = leafletOptions(dragging =F, 
                                 minZoom = 9.5, 
                                 maxZoom = 9.5, 
                                 zoomControl = F)) %>% 
  
  addPolygons(fillColor = ~pal_small(ar_cat),
              weight  = 1.3,
              color = "black", 
              highlightOptions = highlightOptions(weight = 4, 
                                                  color = "red", 
                                                  bringToFront = T, 
                                                  opacity = .5), 
              label = ~labels) %>% 
  
  addLegend(pal = pal_small, values = ~ar_cat, 
            title = "Attack Rate<br><span style='font-size:10px;'>per 10 000") %>% 
  
  addControl(paste0(html = "<span style ='font-size:12px;color:#34495E;'>Chloropleth map of Fulton county at peak of the outbreak: <b>", peak, "</b></span><br><span style ='font-size:11px;color:#34495E;'>regions with less than 1000 habitants"))

```
</center>

At the peak of the outbreak: **`r peak`**, the region with the greatest Attack rate was **`r filter(filter(weekly_zip_ar, ISOweek == peak), ar == max(ar))$case_zip` ** with an Attack rate of `r round(digits = 2, filter(filter(weekly_zip_ar, ISOweek == peak), ar == max(ar))$ar)` `r filter(filter(weekly_zip_ar, ISOweek == peak), ar == max(ar))$ci` per 10 000


#### Chloropleth maps per week and Months

Maps of big regions of Fulton county by Week of Outbreak. 

```{r chloropleth for loop}
#create a function that filters data on month and then plot a chloropleth 
month_week_chloro <- function(data_ar, year_month){
  
  ggplot() +
  
  geom_sf(data = filter(data_ar, month == year_month), 
          
          aes(fill = ar_cat), 
          alpha = .5, 
          color = "black", 
          size=.1) +
  
  scale_fill_manual(values = c("#94d2bd", 
                               "#ffe169", 
                               "#edc531", 
                               "#e85d04", 
                               "#dc2f02", 
                               "#bb3e03")) +
  
  ggtitle(year_month) +
  
  facet_wrap (~ISOweek, nrow = 1) +
  
  theme(panel.background = element_blank(), 
          plot.title = element_text(face = "bold"),
          axis.text = element_blank(), 
          axis.ticks = element_blank(), 
          strip.text = element_text(face="bold"), 
          strip.background = element_rect(color="black", fill="white"), 
          panel.border = element_blank(), 
          axis.title= element_blank(), 
          legend.key = element_blank())
  
}

#for loop the function to create all chloropleth subsetted by month
for (time in(unique(weekly_zip_ar_sf_big$month))) {
  
print(month_week_chloro(weekly_zip_ar_sf_big, year_month = time) )
  
  
}


```









